---
title: "Dow Jones Case Study"
author: "Dylan Rowe"
date: "February 10, 2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
dowdata = read.csv(".../Dow Jones Case Study/dow_jones_index.data")
```

```{r}
dowdata = na.omit(dowdata)
str(dowdata)
summary(dowdata)
```

```{r}
head(dowdata)
```


#We need to clean the data and remove those dollar signs from our variables
```{r}
dowdata$quarter = as.factor(dowdata$quarter)
dowdata$stock = as.factor(dowdata$stock)
dowdata$open = as.numeric(gsub('\\$', '',dowdata$open))
dowdata$high = as.numeric(gsub("\\$", "", dowdata$high))
dowdata$low = as.numeric(gsub("\\$", "", dowdata$low))
dowdata$close = as.numeric(gsub("\\$", "", dowdata$close))
dowdata$next_weeks_open = as.numeric(gsub("\\$", "", dowdata$next_weeks_open))
dowdata$next_weeks_close = as.numeric(gsub("\\$", "", dowdata$next_weeks_close))
dowdata$date = as.Date(dowdata$date, format = "%m/%d/%Y")
dowdata2 = dowdata
```

#Look at a plot of the lags to see if any lags need to be included in our data
```{r}
lag.plot(dowdata$open,6)
lag.plot(dowdata$close,6)
lag.plot(dowdata$low,6)
lag.plot(dowdata$high,6)
```

```{r}
library(dplyr)
dowdata = dowdata %>% group_by(stock) %>% mutate(lag.open = lag(open, n = 1))
dowdata = dowdata %>% group_by(stock) %>% mutate(lag.close = lag(close, n = 1))
dowdata = dowdata %>% group_by(stock) %>% mutate(lag.low = lag(low, n = 1))
dowdata = dowdata %>% group_by(stock) %>% mutate(lag.high = lag(high, n = 1))
dowdata = dowdata %>% group_by(stock) %>% mutate(lag.volume = lag(volume, n = 1))
dowdata = dowdata %>% group_by(stock) %>% 
  mutate(lag.percent_change_volume_over_last_wk 
         = lag(percent_change_volume_over_last_wk, n = 1))
```

#Let's make some histograms to look at the distribution of our data
```{r}
hist(dowdata$open)
hist(dowdata$close)
hist(dowdata$low)
hist(dowdata$high)
```

#Now we will split the data into a test and train set by quarter 1 and quarter 2
```{r}
dowdata_split = split(dowdata, dowdata$quarter)
dow_train = dowdata_split[[1]]
dow_test = dowdata_split[[2]]
```

#We will also split the data on stock in order to run a model for each stock
```{r}
dowtrain_split = split(dow_train, dow_train$stock)
dowtest_split = split(dow_test, dow_test$stock)
```

#Use a for loop to run the 30 stock models using a linear regression and store the MSE predicted on the test set of data
```{r}
#Create an empty vector to store MSE values
MSE = c()
AccLin = c()
n = names(dowtrain_split)
for (i in 1:30) {
  y = dowtest_split[[i]]$percent_change_price
  model <- glm(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[i]])
  model.preds = predict(model, newdata = dowtest_split[[i]])
  MSE[i] = mean((y - model.preds)^2)
  y2 = ifelse(dowtest_split[[i]]$percent_change_price > 0.5,1,0)
  model.class = ifelse(model.preds > 0.5,1,0)
  AccLin[i] = 1 - mean(y2!=model.class)
  #print(summary(model))
}
LinMSE = cbind(n,MSE)
LinAcc = cbind(n,AccLin)
```

#Use a for loop to run a decision tree for the 30 stock companies and store the MSE predicted on the test set of data
```{r}
library(tree)
MSEtree = c()
Acctree = c()
for (i in 1:30) {
  y = dowtest_split[[i]]$percent_change_price
  model <- tree(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open
                + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[i]])
  model.preds = predict(model, newdata = dowtest_split[[i]])
  MSEtree[i] = mean((y - model.preds)^2)
  y2 = ifelse(dowtest_split[[i]]$percent_change_price > 0.5,1,0)
  model.class = ifelse(model.preds > 0.5,1,0)
  Acctree[i] = 1 - mean(y2!=model.class)
}
TreeMSE = cbind(n,MSEtree)
TreeAcc = cbind(n,Acctree)
```

#Use a for loop to run a basic SVM model for our 30 stock companies and store the MSE predicted on the test set of data
```{r}
dowtrain_split2 = dowtrain_split
dowtest_split2 = dowtest_split
library(e1071)
MSEsvm = c()
AccSVM = c()
SVM.Preds = c()
for (i in 1:30) {
  dowtest_split2[[i]][,c(9,17:22)] = scale(dowtest_split2[[i]][,c(9,17:22)])
  dowtrain_split2[[i]][,c(9,17:22)] = scale(dowtrain_split2[[i]][,c(9,17:22)])
  y = scale(dowtest_split2[[i]]$percent_change_price)
  model = svm(percent_change_price ~ lag.high + lag.low + lag.close
              + lag.volume + lag.open + lag.percent_change_volume_over_last_wk, data = dowtrain_split2[[i]], type = "eps-regression", kernel = "radial")
  model.preds = predict(model, newdata = dowtest_split2[[i]])
  MSEsvm[i] = mean((y - model.preds)^2)
  y2 = ifelse(dowtest_split2[[i]]$percent_change_price > 0.5,1,0)
  model.class = ifelse(model.preds > 0.5,1,0)
  AccSVM[i] = 1 - mean(y2!=model.class)
}
SVMmse = cbind(n,MSEsvm)
SVMAcc = cbind(n,AccSVM)
```

#Take the average of the MSE for the 30 stock models ran for each machine learning technique (LM, tree, SVM) to determine the best model on our data
```{r}
LinMean = mean(MSE); LinMean
TreeMean = mean(MSEtree); TreeMean
SVMmean = mean(MSEsvm);SVMmean
LinMeanAcc = mean(AccLin); LinMeanAcc
TreeMeanAcc = mean(Acctree); TreeMeanAcc
SVMmeanAcc = mean(AccSVM);SVMmeanAcc
```

#The lowest average MSE and highest accuracy model is the SVM Radial model and that is the model we will use going forward.

#We will use the last row(week) of the data to predict a single percent change price value for the next week in the future
```{r}
SVM.Preds = c()
for (i in 1:30) {
  dowtest_split2[[i]][,c(9,17:22)] = scale(dowtest_split2[[i]][,c(9,17:22)])
  dowtrain_split2[[i]][,c(9,17:22)] = scale(dowtrain_split2[[i]][,c(9,17:22)])
  y = scale(dowtest_split2[[i]][13,]$percent_change_price)
  model = svm(percent_change_price ~ lag.high + lag.low + lag.close
              + lag.volume + lag.open + lag.percent_change_volume_over_last_wk,
              data = dowtrain_split2[[i]], type = "eps-regression")
  model.preds = predict(model, newdata = dowtest_split2[[i]][13,])
  SVM.Preds[i] = model.preds
}
Pred.SVM = cbind(n,SVM.Preds);Pred.SVM
```


```{r}
Pred.SVM = as.data.frame(Pred.SVM);Pred.SVM
```

#Now we will take the 8 Stocks with the lowest MSE to test and determine the most significant predictors. Those stocks are BAC(4), INTC(14), JPM(16), KO(17), MCD(19), PFE(23), VZ(28), WMT(29).
Below we will create all 8 models but will not look at the output yet.
```{r}
BAC.mod = tree(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open
                + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[4]])
INTC.mod = tree(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open
                + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[14]])
JPM.mod = tree(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open
                + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[16]])
KO.mod = tree(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open
                + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[17]])
MCD.mod = tree(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open
                + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[19]])
PFE.mod = tree(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open
                + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[23]])
VZ.mod = tree(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open
                + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[28]])
WMT.mod = tree(percent_change_price ~ lag.high + lag.low + lag.close + lag.volume + lag.open
                + lag.percent_change_volume_over_last_wk, data = dowtrain_split[[29]])
```

```{r}
par(mfrow = c(2,2))
plot(BAC.mod)
text(BAC.mod)
plot(INTC.mod)
text(INTC.mod)
plot(JPM.mod)
text(JPM.mod)
plot(KO.mod)
text(KO.mod)
plot(MCD.mod)
text(MCD.mod)
plot(PFE.mod)
text(PFE.mod)
plot(VZ.mod)
text(VZ.mod)
plot(WMT.mod)
text(WMT.mod)
```
#All variables appear in decision trees so it would be wise to include all the predictors in model.


#Now we will look at the risk of each stock using the CAPM model. First we will import the data for the SP500 as our reference market.
```{r}
SP500Data = read.csv("C:/Users/Dylan/Documents/MSDA/SEM3/Data Analytics Applications/Dow Jones Case Study/^GSPC.csv", header = TRUE, sep = ",")
head(SP500Data)
```

```{r}
dowstocksplit = split(dowdata, dowdata$stock)
```

```{r}
DataStock = as.data.frame(cbind(1:24))
DataStock = DataStock[-1]
```


#Next we will use a for loop to extract the return from all of our stock variables
```{r}
for (i in 1:30) {
  DataStock[i] = dowstocksplit[[i]]$close
}
colnames(DataStock) = n
DataStock
```

#Now to take the delt of close to get the return
```{r}
library(tseries)
library(quantmod)
AA = Delt(DataStock$AA)
AXP = Delt(DataStock$AXP)
BA = Delt(DataStock$BA)
BAC = Delt(DataStock$BAC)
CAT = Delt(DataStock$CAT)
CSCO = Delt(DataStock$CSCO)
CVX = Delt(DataStock$CVX)
DD = Delt(DataStock$DD)
DIS = Delt(DataStock$DIS)
GE = Delt(DataStock$GE)
HD = Delt(DataStock$HD)
HPQ = Delt(DataStock$HPQ)
IBM = Delt(DataStock$IBM)
INTC = Delt(DataStock$INTC)
JNJ = Delt(DataStock$JNJ)
JPM = Delt(DataStock$JPM)
KO = Delt(DataStock$KO)
KRFT = Delt(DataStock$KRFT)
MCD = Delt(DataStock$MCD)
MMM = Delt(DataStock$MMM)
MRK = Delt(DataStock$MRK)
MSFT = Delt(DataStock$MSFT)
PFE = Delt(DataStock$PFE)
PG = Delt(DataStock$PG)
t = Delt(DataStock$T)
TRV = Delt(DataStock$TRV)
UTX = Delt(DataStock$UTX)
VZ = Delt(DataStock$VZ)
WMT = Delt(DataStock$WMT)
XOM = Delt(DataStock$XOM)
SP500 = na.omit(Delt(SP500Data[,5]))
```

#We will now combine the returns data for all of our stocks into one place
```{r}
ReturnData = cbind(AA, AXP, BA, BAC, CAT, CSCO, CVX, DD, DIS, GE, HD, HPQ, IBM, INTC, JNJ, JPM, KO, KRFT, MCD, MMM, MRK, MSFT, PFE, PG, t, TRV, UTX, VZ, WMT, XOM)
ReturnData = na.omit(ReturnData)
ReturnData = cbind(ReturnData,SP500)
colnames(ReturnData) = c("AA", "AXP", "BA", "BAC", "CAT", "CSCO", "CVX", "DD", "DIS", "GE", "HD", "HPQ", "IBM", "INTC", "JNJ", "JPM", "KO", "KRFT", "MCD", "MMM", "MRK", "MSFT", "PFE", "PG", "t", "TRV", "UTX", "VZ", "WMT", "XOM", "SP500")
```

```{r}
boxplot(ReturnData,main="Expected Return", xlab="Stock Picks", ylab="Return")
```

```{r}
ReturnData = as.data.frame(ReturnData);ReturnData
```


```{r}
Beta = c()
for (i in 1:30) {
  model = lm(ReturnData[,i] ~ SP500, data = ReturnData)
  Beta[i] <- summary(model)$coefficients[2, 1]
}
Beta = cbind(n, Beta)
Beta = as.data.frame(Beta); Beta
```

```{r}
library(car)
library(ggplot2)
Preds = Pred.SVM$SVM.Preds
BetaReturn = cbind(Beta, Preds)
BetaReturn$Beta = as.numeric(as.character(BetaReturn$Beta))
BetaReturn$Preds = as.numeric(as.character(BetaReturn$Preds))
BetaReturn
ggplot(BetaReturn, 
       aes(x = BetaReturn$Beta, y = BetaReturn$Preds, color = "green",
           label = BetaReturn$n)) + 
  geom_point() + 
  geom_text(aes(label = BetaReturn$n), hjust = 1, vjust = 1)
```

